Statistical Inference for inter-arrival times of extreme events in bursty time series

### 2020-03-28 article_springer/article_springer.Rmd
Überarbeitung, Finale Änderung bzgl Reviewer 2
Final changes, adressing Rev2

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  message = FALSE,
  fig.height = 3, out.width = '\\textwidth'
  )
library(CTRE)
library(MittagLeffleR)
library(magrittr)
library(gridExtra)
library(ggplot2)
library(dplyr)
set.seed(12345)
```

@@

In the following we denote with $x_L \in [-\infty, +\infty)$ and $x_R \in (-\infty, +\infty]$) the left and right endpoint of the distribution of $J$. We assume that the $k$-th magnitude $J_k$ occurs at time
$T_k = W_1 + \ldots + W_k$. 
Based on an MRP, we define the Continuous Time Random Exceedance model
(CTRE) as follows:

@@

In this article, we restrict ourselves to the uncoupled case, where $W$ and $J$ are independent. Then the two sequences $X(\ell, n)_{n \in \mathbb N}$ and $T(\ell, n)_{n \in \mathbb N}$ are independent as well. To see why, note that $X(\ell)$ is, in distribution, simply equal to $J - \ell | J > \ell$, independent of any waiting time $W_k$.  
We assume for the rest of the article, that the magnitudes $(J_i)_{i \in \mathbb{N}}$ belong to the the max-domain of attraction of some non-degenerate distribution. This means there exist $a_n>0$ and $d_n \in \mathbb{R}$ such that 
$$ a_n^{-1}(J_1 \vee \ldots \vee J_n -d_n) \Rightarrow A \text{ as } n \rightarrow \infty.$$
Hence, the distribution of $A$ is a generalized extreme value distribution (GEV) which distribution function is given by 
$$ F(x;\xi) = \begin{cases}\exp(-(1+\xi x)^{-1/\xi}) & \xi\neq0 \\ \exp(-\exp(-x)) & \xi = 0\end{cases}.$$ 
The GEV is subdivided into the Gumbel ($\xi=0$), the Weibull ($\xi<0$) and the Fr'echet ($\xi>0$) family of distributions.

@@

*Proof of Theorem:* We interpret the threshold crossing time $T(\ell)$ as the hitting time of the 
underlying CTRM (Continuous Time Random Maxima) or "max-renewal process", 
and then utilize a result by @MeerschaertStoev08. 
The running maximum process is defined as 
$$
M(c) := J_1 \vee \ldots \vee J_{\lfloor c \rfloor},
$$
and since we assume that the $J_k$ have a continuous distribution, there exist 
norming functions $a(c)$ and $d(c)$ such that 
$$
\mathbf P\left[ \frac{M(c) - d(c)}{a(c)} \le \ell^* \right] 
\longrightarrow F(\ell^*), \quad c \to \infty
$$
where $F$ is a generalized extreme value distribution, and $\ell^*$ is any 
value from the support of $F$. 
The CTRM process is then defined via 
$$
V(t) = M(N(t)), \quad t \ge 0
$$
where $N(t)$ is the renewal process associated with the waiting times 
$W_k$: 
$$
N(t) = \max\{n: W_1 + \ldots + W_n \le t\}.
$$
Now a key observation is that
$$
T(\ell) = \inf\{t: V(t) > \ell\}, 
$$
and that

@@

The classical POT approach is based on the fact that exceedances above a high threshold are asymptotically GPD distributed, hence a GPD distribution is fitted to the exceedances above a high enough threshold. Due to the Theorem in the last Section we know that for high thresholds the times between exceedances are asymptotically Mittag-Leffler distributed in case of heavy-tailed waiting times between the observations. Hence, before we talk about inference for the exceedances in the next section, we discuss inference for Mittag-Leffler distributions.

Historically, the first method proposed for the estimation of the Mittag-Leffler distribution parameters was the fractional moment estimator by @kozubowski2001. Unlike the first moments, the fractional moments of order $p$ for $p<\beta$ exist and are tractable. One drawback of this method is that constant priors for the tail parameter are needed for the calculation of the estimates. @Cahoy2010 proposed a moment estimator of the log-transformed data, which does not require any prior. Furthermore, they performed simulation studies illustrating that the log-moment outperforms the fractional moment estimator with respect to bias and root mean squared error (RMSE).

Due to the form of the density function of a Mittag-Leffler distribution there exists no closed form for the Maximum Likelihood estimator (MLE). In the R Package `MittagLeffleR` [@MittagLeffleR], Maximum Likelihood estimation is implemented via numerical optimization. The MLE slightly outperforms the log-moment estimator regarding bias and RMSE for big enough sample sizes, but is extremely computationally intensive. Figure \ref{fig:MSE} shows that both estimators perform well, even for small sample sizes.

```{r MSE, echo=FALSE,warning=FALSE, fig.height=7,fig.width=8,fig.align="center", fig.cap="\\label{fig:MSE} MSE for the estimation of tail (left column) and scale (right column) parameters via log-moment estimator and MLE of a Mittag-Leffler sample with varying sample size n=25, 100, 200, varying tails on the x-axis and fixed scale equal to one.", out.width = '90%'}
load(here::here("article_springer", "article_springer_files", "ML_estim_simu","tbl_ges.RData"))
tbl_ges %>% 
  filter(n == "25" || n == "100" || n=="200") %>%
  filter(true.scale == "1") %>% 
  ggplot(aes(true.tail, mse, colour = method, group=method)) + 
  geom_line() + geom_point() + 
  facet_wrap(c("n","param"), scales = "free", ncol = 2) + 
  scale_color_manual(values=c("red", "blue"))
```

Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator [@hill1975simple] based on the $r+1$ upper order statistics
\begin{align}\label{eq:Hill}
H_{r,n}=\left[ \frac{1}{r} \sum_{i=0}^{r} \log \frac{X_{(i)}}{X_{(r+1)}}\right]^{-1}, 
\end{align}
where $X_{(1)} \geq X_{(2)} \geq \ldots \geq X_{(n)}$ denote the order statistics in decreasing order of a sequence $X_1,\ldots,X_n$. However, these methods are statistically less efficient since they only use a portion of the information contained in the data, as mentioned by @kozubowski2001. Moreover, the Hill estimator requires a tuning parameter, denoted with $r$ in \eqref{eq:Hill}. Additionaly, the estimator only performs reliably well for distributions close to Pareto [see @Resnick97]. Figure \ref{fig:Hillplots} shows Hill plots for Mittag-Leffler simulated data, with varying sample sizes and tails. To deduce the correct tail parameter estimates from these plots is virtually impossible. To estimate the tail parameter of a Mittag-Leffler distribution with the Hill estimator, the sample size has to be much larger and one has to choose a large tuning parameter $r$. In case that only events above a high threshold are recorded, even $n=1000$ events are unrealistic. Furthermore the Hill estimator of course completely fails if the inter-arrival times are exponentially distributed and not heavy-tailed. We will come back to the problems with the Hill estimtator in Section \ref{Simulationstudy}.

```{r Hillplots, echo=FALSE,warning=FALSE, fig.align="center", fig.cap="\\label{fig:Hillplots} Hillplots for simulated Mittag-Leffler data with true tail 0.8 and sample size 200 (left) and 1000 (right), with number of upper order statics r on which the Hill estimator is based on the x-axis. ", out.width = '90%'}
hill_estim<-function(k,data){
  n<-length(data)
  data<-sort(data)
  data_thinned<-data[(n-k):n]
  r<-1/k * sum(log(data_thinned/data_thinned[1]))
  return(1/r)
}

hill_estim_plot<-function(k_min,data,k_max=n,true_tail){
  n<-length(data)
  x_values<-k_min:k_max
  y_values<-sapply(x_values,hill_estim,data=data)
  df<-data.frame(x=x_values,y=y_values)
  ggplot2::ggplot(data=df,aes(x_values,y_values))+ geom_line(color="blue")+xlab("Order Statistics") + ylab("Hill-Estimator of Tail Index") +
  coord_cartesian(ylim=c(0,1.5)) + geom_hline(yintercept=true_tail, linetype="dashed", color = "red", size=0.5)
}

data_ml_1<-rml(n=200,tail=0.8,scale=1)
p1<-hill_estim_plot(k_min=30,data=data_ml_1,true_tail=0.8) +
  ggtitle("Hill plot, Mittag-Leffler data, n=200")

data_ml_2<-rml(n=1000,tail=0.8,scale=1)
p2<-hill_estim_plot(k_min=30,data=data_ml_2,true_tail=0.8) +
  ggtitle("Hill plot, Mittag-Leffler data, n=1000")

@@

Since the exponential distribution is nested in the Mittag-Leffler family of 
distributions, a Likelihood-ratio Test (LRT) seems to be an appropriate way to choose 
between a model with exponential and Mittag-Leffler inter-exceedance times.
Although the two models are nested, the asymptotic distribution is not 
$\chi^2_1$-distributed, and Wilk's Theorem does not hold: under $H_0$, the 
parameter $\beta$ of Mittag-Leffler distribution is equal to $1$, and hence lies
on the boundary of the parameter space $(0,1]$. 
Instead, a valid approach is a bootstrapped Likelihood-ratio test [see e.g. @davison1997bootstrap]. 
Figure \ref{Fig:LRT} displays the (simulated) power for the bootstrapped LRT for Mittag-Leffler distributions with varying tail parameters based on 1000 simulation runs. As expected, the power decreases for tail parameters close to one, since the Mittag-Leffler distribution converges as $\beta \uparrow 1$ to an exponential distribution; it becomes hard to differentiate a Mittag-Lefller distribution from an exponential.

@@

2. Choose e.g. $K_0 = 5$, and for each $k$ ranging from $K_0 +1$ to $K + 1$: 
    a) extract the set $\mathcal T_k$ of exceedance times of between the magnitudes exceeding the threshold $X_{(k)}$
    b) fit a Mittag-Leffler distribution to $\mathcal T_k$, resulting in the parameter
    estimates $\hat\beta_k$ and $\hat \sigma_k$.

@@

To clarify Step 4: Recall that by the theorem, $\sigma_k / b(1/p_{\ell})$ is expected to stabilize 
around a constant as $k \downarrow K_0$. Since $b$ is regularly varying with parameter
$1/\beta$, we have $b(1/p_{\ell}) = p_{\ell}^{-1/\beta} / L(1/p_{\ell})$ for some slowly varying function $L$. 
Approximating $p_{\ell} \approx \hat p_{\ell} := k / n$, we have 
$$\text{const} \approx \sigma_k / b(1/p_{\ell}) = \sigma_k \times p_{\ell}^{1/\beta} L(1/p_{\ell}) \approx \sigma_k \times k^{1/\beta} n^{-1/\beta} L(n/k)$$
Assuming that the variation of $L(n/k)$ is minor, we can hence see that 
$\sigma_k k^{1/\beta}$ stabilizes.

@@

To test our inference method via stability plots, we have simulated $m=100$ times $n=10000$ independent waiting time and magnitude pairs $(W_k, J_k)$ for waiting times that follow

@@

+ rows correspond to cases (i), (ii) and (iii), and
+ columns correspond to estimators (log-moment estimator, maximum-likelihood, and Hill estimator).

We plot the tail parameter estimates $\hat \beta(k)$ against $k$ 
for each of the $m=100$ simulation runs. 
Thin grey lines represent individual simulation runs, 
and the thicker black line is their mean. 
Recall that $k$ is the index of the order statistics of $J_k$ at which the 
threshold $\ell$ is placed. 
Note that the Hill estimator is based on the waiting times, not on the 
inter-exceedance times, meaning that the estimation is based on more data. 
If one were to base the Hill-estimation only on the inter-exceedance times, 
the sample size would be too small and it would be impossible to deduce an 
estimate for the tail parameter from the plots (compare Figure \ref{fig:Hillplots}).

In the right-hand Hill-estimator column, the $k$ on the $x$-axis corresponds to
$r$ from \eqref{eq:Hill}.
The Hill estimator appears to be nicely unbiased on Pareto data,
but compared to the other two estimators, it has a higher variance, 
as can be seen from the wider spread of the simulated curves. 
This high variance means that that a single estimate is less reliable compared
to the other two estimators. 
This effect is amplified if one only has a thresholded, and thus smaller dataset
available, to the extent that the Hill estimates become severely flawed
(see Section \ref{sec:ML}).

Moreover, the bottom in \ref{Fig:TailSimu} shows clearly that the log-moment 
and maximum-likelihod estimators generalize to the exponential case ($\beta = 1$), 
whereas the Hill-estimator does not.

Finally, the performance of Log-Moment and Maximum-Likelihood estimators for the
scale parameter is shown in figure \ref{Fig:ScaleSimu}. 
Since there is no Hill-estimate of a scale parameter, there are only two columns. 
Both estimators show good performance, with a slight advantage for the 
Maximum-Likelihood estimator. This advantage is paid for by a higher computational 
cost.

```{r TailSimuplots, message=FALSE,echo=FALSE,fig.height=7,fig.align="center",fig.cap="\\label{Fig:TailSimu} Stability Plots for m=100 simulation runs for stable distributed waiting times with a tail parameter of 0.8 (top row), Pareto distributed waiting times with tail parameter 0.8 (middle row) and exponentially distributed waiting times (lower row). Left column: log-moment estimator, middle column: MLE, right column: Hill estimator. The grey thin lines are the stability plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true tail parameter. ", fig.show='hold', out.width = '100%'}

@@

#Hill estimator
hill_Ex1_tail <-
  ggplot() + geom_line(data = tbl_estimates_tail08_hill_Ex1,
                       aes(x = k, y = value, group = rep),
                       colour = "lightgrey") +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(data=means.tbl_hill_Ex1,aes(x=k, y=mw),colour="black") + 
  ylab("Estimated tail") +
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+ xlab("r")

@@

#Hill estimator
hill_Ex2_tail <-
  ggplot() + geom_line(aes(x = k, y = value, group = rep),
                       colour = "lightgrey",
                       data = tbl_estimates_Ex2_Pareto_tail08_hill_times) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x = k, y = mw), colour = "black", data = means.tbl_hill_times_Ex2) + 
  ylab("Estimated tail") +
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+ xlab("r")

@@

#Hill estimator
hill_Ex3_tail <-
  ggplot() + geom_line(aes(x = k, y = value, group = rep),
                       colour = "lightgrey",
                       data = tbl_estimates_Example3_Exp_hill_times) +
  coord_cartesian(ylim = c(0.5, 10)) +
  geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_hill_times_Ex3) + 
  ylab("Estimated tail") + 
  geom_hline(yintercept=1, linetype="dashed", color = "red")+ xlab("r")

grid.arrange(
  arrangeGrob(logMom_Ex1_tail,logMom_Ex2_tail,logMom_Ex3_tail,top="log-Moment"),
  arrangeGrob(mle_Ex1_tail, mle_Ex2_tail, mle_Ex3_tail,top="MLE"),
  arrangeGrob(hill_Ex1_tail,hill_Ex2_tail,hill_Ex3_tail,top="Hill"),
  ncol=3
)
```

```{r ScaleSimuplots, message=FALSE,echo=FALSE,fig.height=7,fig.align="center",fig.cap="\\label{Fig:ScaleSimu} Stability Plots for m=100 simulation runs for stable distributed waiting times with a tail parameter of 0.8 (top row), Pareto distributed waiting times with tail parameter 0.8 (middle row) and exponentially distributed waiting times (lower row). Left column: log-moment estimator, right column: MLE. The grey thin lines are the stability plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true tail parameter.", fig.show='hold', out.width = '90%'}
#Scale Parameter

@@

Before we apply the POT approach described in Section 5 to the solar flare data, we first have to check if all model assumptions are fulfilled. The CTRE model is based on three main assumptions, which are repeated below. For each assumption, we suggest one means of checking if it holds:

@@

: After removing the "noise observations" below the smallest threshold 
  $\ell_0$, the pair sequence $(T(\ell_0, i), X(\ell_0,i))$ is i.i.d. 
  An indication if this is true is given 
  by an auto-correlation plot for the logarithms (to ensure finite moments) 
  of the two time series.

@@

: Apply a cutoff at the lowest threshold $\ell_0$, 
  extract the threshold crossing times, 
  and create a QQ Plot for the Mittag-Leffler distribution. 
  Use a log-moment estimate of the tail parameter for the theoretical / population 
  quantiles of the plot.

Figures \ref{fig:flare-diagnostics-1}, \ref{fig:flare-diagnostics-2} and \ref{fig:flare-diagnostics-3} show the diagnostic plots for a minimum threshold chosen at the 200th order statistic. There is some residual autocorrelation for the sequence of threshold exceedance times that is not accounted for by the CTRE model.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  message = FALSE,
  fig.height = 3, out.width = '\\textwidth'
  )
library(CTRE)
library(MittagLeffleR)
library(magrittr)
library(gridExtra)
library(ggplot2)
library(dplyr)
library(forecast)
set.seed(12345)
```

@@

In the following we denote with $x_L \in [-\infty, +\infty)$ and $x_R \in (-\infty, +\infty]$ the left and right endpoint of the distribution of $J$. We assume that the $k$-th magnitude $J_k$ occurs at time
$T_k = W_1 + \ldots + W_k$. 
Based on an MRP, we define the Continuous Time Random Exceedance model
(CTRE) as follows:

@@

In this article, we restrict ourselves to the uncoupled case, where $W$ and $J$ are independent. Then the two sequences $X(\ell, n)_{n \in \mathbb N}$ and $T(\ell, n)_{n \in \mathbb N}$ are independent as well. To see why, note that $X(\ell)$ is, in distribution, simply equal to $J - \ell | J > \ell$, independent of any waiting time $W_k$.  
We assume for the rest of the article, that the magnitudes $(J_i)_{i \in \mathbb{N}}$ belong to the the max-domain of attraction of some non-degenerate distribution. This means there exist $a_n>0$ and $d_n \in \mathbb{R}$ such that 
\begin{align} \label{assumptionJs}
a_n^{-1}(J_1 \vee \ldots \vee J_n -d_n) \Rightarrow A \text{ as } n \rightarrow \infty.
\end{align}
Hence, the distribution of $A$ is a generalized extreme value distribution (GEV) which distribution function is given by 
$$ F(x;\xi) = \begin{cases}\exp(-(1+\xi x)^{-1/\xi}) & \xi\neq0 \\ \exp(-\exp(-x)) & \xi = 0\end{cases}.$$ 
The GEV is subdivided into the Gumbel ($\xi=0$), the Weibull ($\xi<0$) and the Fréchet ($\xi>0$) family of distributions.

@@

*Proof of Theorem:* We interpret the threshold crossing time $T(\ell)$ as the hitting time of the 
underlying CTRM (Continuous Time Random Maxima) or "max-renewal process", 
and then utilize a result by @MeerschaertStoev08. 
The running maximum process is defined as 
$$
M(c) := J_1 \vee \ldots \vee J_{\lfloor c \rfloor},
$$
and since we assumed that the $J_k$ our out of the domain of attraction of a max-stable distribution $F$ (see \eqref{assumptionJs}), there exist 
norming functions $a(c)$ and $d(c)$ such that 
$$
\mathbf P\left[ \frac{M(c) - d(c)}{a(c)} \le \ell^* \right] 
\longrightarrow F(\ell^*), \quad c \to \infty
$$
where $F$ is a generalized extreme value distribution, and $\ell^*$ is any 
value from the support of $F$. 
The CTRM process is then defined via 
$$
V(t) = M(N(t)), \quad t \ge 0
$$
where $N(t)$ is the renewal process associated with the waiting times 
$W_k$: 
$$
N(t) = \max\{n: W_1 + \ldots + W_n \le t\}.
$$
Now a key observation is that
$$
T(\ell) = \inf\{t: V(t) > \ell\}, 
$$
and that

@@

The classical POT approach is based on the fact that exceedances above a high threshold are asymptotically GPD distributed, hence a GPD distribution is fitted to the exceedances above a high enough threshold. This approach was well-established a long time ago and is standard in extreme value theory nowadays. Hence we just refer here to the according literature (@smith1984threshold, @leadbetter1991basis, @davison1990models), a nice introduction can be found in @beirlantBook Ch.5.3, @embrechts2013modelling Ch. 6.5. or @ColesBook Ch.4. Hence, we focus in this article mainly on the inter-exceedance times. Due to the Theorem in the last Section we know that for high thresholds the times between exceedances are asymptotically Mittag-Leffler distributed in case of heavy-tailed waiting times between the observations. In this article we propose to proceed with inter-exceedance times in a similar fashion as with the magnitudes in the classical POT approach, more precisely to fit a Mittag-Leffler distribution via stability plots to the inter-exceedance times. Hence, before we talk about inference for the inter-exceedance times in the next section, we first discuss inference for Mittag-Leffler distributions.

Historically, the first method proposed for the estimation of the Mittag-Leffler distribution parameters was the fractional moment estimator by @kozubowski2001. Unlike the first moments, the fractional moments of order $p$ for $p<\beta$ exist and are tractable. One drawback of this method is that constant priors for the tail parameter are needed for the calculation of the estimates. @Cahoy2010 proposed a moment estimator of the log-transformed data, which does not require any prior. Furthermore, they performed simulation studies illustrating that the log-Moment outperforms the fractional moment estimator with respect to bias and root mean squared error (RMSE).

Due to the form of the density function of a Mittag-Leffler distribution there exists no closed form for the Maximum Likelihood estimator (MLE). In the R Package `MittagLeffleR` [@MittagLeffleR], Maximum Likelihood estimation is implemented via numerical optimization. The MLE slightly outperforms the log-Moment estimator regarding bias and RMSE for big enough sample sizes, but is extremely computationally intensive. Figure \ref{fig:MSE} shows that both estimators perform well, even for small sample sizes.

```{r MSE, echo=FALSE,warning=FALSE, fig.height=7,fig.width=8,fig.align="center", fig.cap="\\label{fig:MSE} MSE for the estimation of tail (left column) and scale (right column) parameters via log-Moment estimator and MLE of a Mittag-Leffler sample with varying sample size n=25, 100, 200, varying tails on the x-axis and fixed scale equal to one, based on 5000 simulation runs.", out.width = '90%'}
load(here::here("article_springer", "article_springer_files", "ML_estim_simu","tbl_ML_estim.RData"))
tbl_ML_estim %>% filter(method=="logMom" | method=="ML")%>%
  filter(n == "25" || n == "100" || n=="200") %>%
  filter(true.scale == "1") %>% 
  ggplot(aes(true.tail, rmse, colour = method, group=method)) + 
  geom_line() + geom_point() + 
  facet_wrap(c("n","param"), scales = "free", ncol = 2) + 
  scale_color_manual(values=c("red", "blue"))
```

Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators for the estimation of the parameter $\beta$ such as the Hill estimator [@hill1975simple] based on the $r+1$ upper order statistics
\begin{align}\label{eq:Hill}
H_{r,n}=\left[ \frac{1}{r} \sum_{i=0}^{r} \log \frac{X_{(i)}}{X_{(r+1)}}\right]^{-1}, 
\end{align}
where $X_{(1)} \geq X_{(2)} \geq \ldots \geq X_{(n)}$ denote the order statistics in decreasing order of a sequence $X_1,\ldots,X_n$. The Hill estimator only performs reliably well for distributions close to Pareto [see @Resnick97]. It heavily depends on the choice of $r$ in \eqref{eq:Hill}, hence on the number of order statistics entering into \eqref{eq:Hill}.  Figure \ref{fig:Hillplots} shows Hill plots for Mittag-Leffler simulated data, with varying sample sizes and tails. To deduce the correct tail parameter estimates from these plots is virtually impossible. To estimate the tail parameter of a Mittag-Leffler distribution with the Hill estimator, the sample size has to be much larger and one has to choose a large tuning parameter $r$. In case that only events above a high threshold are recorded, $n=10000$ events are unrealistic. There is a vast literature on bias-reduced versions of the Hill estimator and on the optimal choice of $r$, mainly based on imposed second-order regular variation conditions [see e.g. @beirlantBook, Ch.4.4 and 4.7 and references in it]. Also, there are lot of other tail-index estimators not taken under consideration in our article, since this would go beyond the scope of our article. However, we know from the Theorem in Section \ref{sec:scaling} that the inter-exceedance times are asymptotically Mittag-Leffler distributed. So all these general tail (and scale) estimation methods are statistically less efficient since they only use a portion of the information contained in the data, as mentioned by @kozubowski2001. Furthermore, the Hill estimator of course completely fails, if the inter-arrival times are exponentially distributed and not heavy-tailed. Hence, we propose to use the log-Moment estimator or the MLE for Mittag-Leffler distributions.

```{r Hillplots, echo=FALSE,warning=FALSE, fig.align="center", fig.cap="\\label{fig:Hillplots} Hillplots for simulated Mittag-Leffler data with true tail 0.8 and sample size 200 (left) and 10000 (right), with number of upper order statics r on which the Hill estimator is based on the x-axis. ", out.width = '90%'}
hill_estim<-function(k,data){
  n<-length(data)
  data<-sort(data)
  data_thinned<-data[(n-k):n]
  r<-1/k * sum(log(data_thinned/data_thinned[1]))
  return(1/r)
}

set.seed(1234)
hill_estim_plot<-function(k_min,data,k_max=n,true_tail){
  n<-length(data)
  x_values<-k_min:k_max
  y_values<-sapply(x_values,hill_estim,data=data)
  df<-data.frame(x=x_values,y=y_values)
  ggplot2::ggplot(data=df,aes(x_values,y_values))+ geom_line(color="blue")+xlab("Order Statistics") + ylab("Hill-Estimator of Tail Index") +
  coord_cartesian(ylim=c(0,1.5)) + geom_hline(yintercept=true_tail, linetype="dashed", color = "red", size=0.5)
}
data_ml_1<-rml(n=200,tail=0.8,scale=1)
p1<-hill_estim_plot(k_min=30,data=data_ml_1,true_tail=0.8) +
  ggtitle("Hill plot, ML-data, n=200")

data_ml_2<-rml(n=10000,tail=0.8,scale=1)
p2<-hill_estim_plot(k_min=30,k_max=200,data=data_ml_2,true_tail=0.8) +
  ggtitle("Hill plot, ML-data, n=10000")

@@

Since the exponential distribution is nested in the Mittag-Leffler family of 
distributions, a Likelihood-ratio Test (LRT) seems to be an appropriate way to choose 
between a model with exponential and Mittag-Leffler inter-exceedance times.
Although the two models are nested, the asymptotic distribution is not 
$\chi^2_1$-distributed, and Wilk's Theorem does not hold: under $H_0$, the 
parameter $\beta$ of Mittag-Leffler distribution is equal to $1$, and hence lies
on the boundary of the parameter space $(0,1]$. 
Instead, a valid approach is a bootstrapped Likelihood-ratio test [see e.g. @davison1997bootstrap]. 
Figure \ref{Fig:LRT} displays the (simulated) power for the bootstrapped LRT for Mittag-Leffler distributions with varying tail parameters based on 1000 simulation runs. As expected, the power decreases for tail parameters close to one, since the Mittag-Leffler distribution converges as $\beta \uparrow 1$ to an exponential distribution; it becomes hard to differentiate a Mittag-Leffler distribution from an exponential.

@@

2. Choose a minimum number of exceedances $K_0\leq K$, e.g. $K_0 = 5$, and for each $k$ ranging from $K_0$ to $K$: 
    a) extract the set $\mathcal T_k$ of exceedance times between the magnitudes exceeding the threshold $X_{(k)}$
    b) fit a Mittag-Leffler distribution to $\mathcal T_k$, resulting in the parameter
    estimates $\hat\beta_k$ and $\hat \sigma_k$.

@@

To clarify Step 4: Recall that by the theorem, $\sigma_k / b(1/p_{\ell})$ is expected to stabilize 
around a constant as $k$ gets smaller. Since $b$ is regularly varying with parameter
$1/\beta$, we have $b(1/p_{\ell}) = p_{\ell}^{-1/\beta} / L(1/p_{\ell})$ for some slowly varying function $L$. 
Approximating $p_{\ell}$ by $\hat p_{\ell} := k / n$, we have 
$$\text{const} \approx \sigma_k / b(1/p_{\ell}) = \sigma_k \times p_{\ell}^{1/\beta} L(1/p_{\ell}) \approx \sigma_k \times k^{1/\beta} n^{-1/\beta} L(n/k)$$
Assuming that the variation of $L(n/k)$ is minor, we can hence see that 
$\sigma_k k^{1/\beta}$ stabilizes.

**Remark**:

: We approximated $p_{\ell}$, the probability that an event is larger than $l$, by its relative frequency. One could also approximate this tail probability by the GPD distribution fitted to the exceedances.

@@

To test our inference method via stability plots, we have simulated $m=100$ datasets with $n=10000$ independent waiting time and magnitude pairs $(W_k, J_k)$ for waiting times that follow

@@

+ rows correspond to cases (i), (ii) and (iii), and
+ columns correspond to estimators (log-Moment estimator, Maximum-Likelihood).

We plot the tail parameter estimates $\hat \beta(k)$ against $k$ 
for each of the $m=100$ simulation runs. 
Thin grey lines represent individual simulation runs, 
and the thicker black line is their mean. 
Recall that $k$ is the index of the order statistics of $J_k$ at which the 
threshold $\ell$ is placed. 
<!-- Note that the Hill estimator is based on the waiting times, not on the  -->
<!-- inter-exceedance times, meaning that the estimation is based on more data.  -->
<!-- If one were to base the Hill-estimation only on the inter-exceedance times,  -->
<!-- the sample size would be too small and it would be impossible to deduce an  -->
<!-- estimate for the tail parameter from the plots (compare Figure \ref{fig:Hillplots}).  -->

<!-- In the right-hand Hill-estimator column, the $k$ on the $x$-axis corresponds to -->
<!-- $r$ from \eqref{eq:Hill}. -->
<!-- The Hill estimator appears to be nicely unbiased on Pareto data, -->
<!-- but compared to the other two estimators, it has a higher variance,  -->
<!-- as can be seen from the wider spread of the simulated curves.  -->
<!-- This high variance means that that a single estimate is less reliable compared -->
<!-- to the other two estimators.  -->
<!-- This effect is amplified if one only has a thresholded, and thus smaller dataset -->
<!-- available, to the extent that the Hill estimates become severely flawed -->
<!-- (see Section \ref{sec:ML}).  -->

The performance of log-Moment and Maximum-Likelihood estimators for the tail resp. the scale parameter is shown in Figure \ref{Fig:TailSimu} resp. \ref{Fig:ScaleSimu}. Both estimators show good performance, with a slight advantage for the Maximum-Likelihood estimator. This advantage is paid for by a higher computational cost. Moreover, the bottom row in Figure \ref{Fig:TailSimu} shows clearly that the log-Moment and Maximum-Likelihood estimators generalize to the exponential case ($\beta = 1$) whereas a general tail estimator like the Hill-estimator would not.

<!-- Since there is no Hill-estimate of a scale parameter, there are only two columns.  -->

```{r TailSimuplots, message=FALSE,echo=FALSE,fig.height=7,fig.align="center",fig.cap="\\label{Fig:TailSimu} Stability Plots for m=100 simulation runs for stable distributed waiting times with a tail parameter of 0.8 (top row), Pareto distributed waiting times with tail parameter 0.8 (middle row) and exponentially distributed waiting times (lower row). Left column: log-Moment estimator, right column: MLE. The grey thin lines are the stability plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true tail parameter.", fig.show='hold', out.width = '100%'}

@@

#Hill estimator
# hill_Ex1_tail <-
#   ggplot() + geom_line(data = tbl_estimates_tail08_hill_Ex1,
#                        aes(x = k, y = value, group = rep),
#                        colour = "lightgrey") +
#   coord_cartesian(ylim = c(0.5, 1.5)) +
#   geom_line(data=means.tbl_hill_Ex1,aes(x=k, y=mw),colour="black") + 
#   ylab("Estimated tail") +
#   geom_hline(yintercept=0.8, linetype="dashed", color = "red")+ xlab("r")

@@

# #Hill estimator
# hill_Ex2_tail <-
#   ggplot() + geom_line(aes(x = k, y = value, group = rep),
#                        colour = "lightgrey",
#                        data = tbl_estimates_Ex2_Pareto_tail08_hill_times) +
#   coord_cartesian(ylim = c(0.5, 1.5)) +
#   geom_line(aes(x = k, y = mw), colour = "black", data = means.tbl_hill_times_Ex2) + 
#   ylab("Estimated tail") +
#   geom_hline(yintercept=0.8, linetype="dashed", color = "red")+ xlab("r")

@@

# #Hill estimator
# hill_Ex3_tail <-
#   ggplot() + geom_line(aes(x = k, y = value, group = rep),
#                        colour = "lightgrey",
#                        data = tbl_estimates_Example3_Exp_hill_times) +
#   coord_cartesian(ylim = c(0.5, 10)) +
#   geom_line(aes(x=k, y=mw),colour="black",data=means.tbl_hill_times_Ex3) + 
#   ylab("Estimated tail") + 
#   geom_hline(yintercept=1, linetype="dashed", color = "red")+ xlab("r")
# 
grid.arrange(
  arrangeGrob(logMom_Ex1_tail,logMom_Ex2_tail,logMom_Ex3_tail,top="log-Moment"),
  arrangeGrob(mle_Ex1_tail, mle_Ex2_tail, mle_Ex3_tail,top="MLE"),
  # arrangeGrob(hill_Ex1_tail,hill_Ex2_tail,hill_Ex3_tail,top="Hill"),
  ncol=2
)
```

```{r ScaleSimuplots, message=FALSE,echo=FALSE,fig.height=7,fig.align="center",fig.cap="\\label{Fig:ScaleSimu} Stability Plots for m=100 simulation runs for stable distributed waiting times with a tail parameter of 0.8 (top row), Pareto distributed waiting times with tail parameter 0.8 (middle row) and exponentially distributed waiting times (lower row). Left column: log-Moment estimator, right column: MLE. The grey thin lines are the stability plots for the different simulation runs and the dark lines are their means. The red dotted line shows the true scale parameter.", fig.show='hold', out.width = '100%'}
#Scale Parameter

@@

Before we apply the approach described in Section 5 to the solar flare data, we first have to check if all model assumptions are fulfilled. The CTRE model is based on three main assumptions, which are repeated below. For each assumption, we suggest one means of checking if it holds:

@@

: After removing the "noise observations" below the smallest threshold 
  $\ell_0$, the pair sequence $(T(\ell_0, i), X(\ell_0,i))$ is i.i.d. 
  An indication if this is true is given by an auto-correlation plot. Since we   are expecting the inter-exceedance times to be Mittag-Leffler distributed and hence to have infinite mean but finite log-moments, we first take the logarithms of the times.

@@

: Apply a cutoff at the lowest threshold $\ell_0$, 
  extract the threshold crossing times, 
  and create a QQ Plot for the Mittag-Leffler distribution. 
  Use a Log-moment estimate of the tail parameter for the theoretical / population 
  quantiles of the plot.

: The ACF plots of course can just give an indication whether there are dependencies, since they actually just measure linear dependencies. Furthermore, if one calculates the ACF for the logarithmic inter-exceedance times, the ACF indicates on the original scale a multiplicative dependence.

Figures \ref{fig:flare-diagnostics-1}, \ref{fig:flare-diagnostics-2} and \ref{fig:flare-diagnostics-3} show the diagnostic plots for a minimum threshold chosen at the 100th order statistic. There is some residual autocorrelation for the sequence of threshold exceedance times that is not accounted for by the CTRE model.
